Estimating Predictive Rate-Distortion Curves via Neural Variational Inference

### 2019-05-03 paper.tex
Updates from Overleaf

\DeclarePairedDelimiterX{\infdivx}[2]{(}{)}{%
  #1\;\delimsize\|\;#2%
}
\newcommand{\infdiv}{D\infdivx}
\newcommand{\future}{\overrightarrow{X}}
\newcommand{\past}{\overleftarrow{X}}
\newcommand{\key}{\textbf}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\newcommand{\finitefuture}{\overrightarrow{X}^{M}}
\newcommand{\finitepast}{^{-M}\overleftarrow{X}}%_{-M\dots -1}}

@@

This motivates the study of models that try to extract those key features of past observations that are most relevant to predicting the future.
%This results in a tradeoff between the cost of representing past observations and predicting the future well.
%This tradeoff has been formalized and computationally studied in a range of prior studies.
A general information-theoretic framework for this problem is provided by Predictive Rate-Distortion~~\citep{still-bottleneck-2014,marzen-predictive-2016}, also known as the Past-Future Information Bottleneck~\citep{creutzig-past-future-2009}.
The Predictive Rate-Distortion tradeoff seeks to find an encoding of past observations that is maximally informative about future observations, while satisfying a bound on the amount of information that has to be stored.
More formally, this framework trades off prediction loss in the future, formalized as cross-entropy, with the cost of representing the past, formalized as the mutual information between the past observations and the compressed representations of the past.
Due to its information-theoretic nature, this framework is extremely general and applies to processes across vastly different domains.
It has been applied to linear dynamical systems \citep{creutzig-past-future-2009,amir2015past}, but is equally meaningful for discrete dynamical systems~\citep{marzen-predictive-2016}.
For biological systems that make predictions about their environment, this corresponds to placing an information-theoretic constraint on the computational cost used for conditioning actions on past observations~\citep{genewein2015bounded}.

>>>>

\DeclarePairedDelimiterX{\infdivx}[2]{(}{)}{%
  #1\;\delimsize\|\;#2%
}
\newcommand{\infdiv}{D\infdivx}
\newcommand{\future}{\overrightarrow{X}}
\newcommand{\past}{\overleftarrow{X}}
\newcommand{\key}{\textbf}
\DeclarePairedDelimiter{\norm}{\lVert}{\rVert}
\newcommand{\finitefuture}{\stackrel{\rightarrow \scriptscriptstyle{M}}{X}}
\newcommand{\finitepast}{\stackrel{\scriptscriptstyle{M}\leftarrow}{X}}%_{-M\dots -1}}

@@

This motivates the study of models that try to extract those key features of past observations that are most relevant to predicting the future.
%This results in a tradeoff between the cost of representing past observations and predicting the future well.
%This tradeoff has been formalized and computationally studied in a range of prior studies.
A general information-theoretic framework for this problem is provided by Predictive Rate-Distortion~~\citep{still-information-2014,marzen-predictive-2016}, also known as the Past-Future Information Bottleneck~\citep{creutzig-past-future-2009}.
The Predictive Rate-Distortion tradeoff seeks to find an encoding of past observations that is maximally informative about future observations, while satisfying a bound on the amount of information that has to be stored.
More formally, this framework trades off prediction loss in the future, formalized as cross-entropy, with the cost of representing the past, formalized as the mutual information between the past observations and the compressed representations of the past.
Due to its information-theoretic nature, this framework is extremely general and applies to processes across vastly different domains.
It has been applied to linear dynamical systems \citep{creutzig-past-future-2009,amir2015past}, but is equally meaningful for discrete dynamical systems~\citep{marzen-predictive-2016}.
For biological systems that make predictions about their environment, this corresponds to placing an information-theoretic constraint on the computational cost used for conditioning actions on past observations~\citep{genewein2015bounded}.
